Predictive modeling with cross validation

### 2017-08-17 model fitting/cross_validation.Rmd
finished cross validation, wrote function to take data set, fit model and extract predicted survival probabilities

```{r, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

@@

##### Fitting model with all significant variables (univariate analysis p-value < 0.01), select variables using stepwise functions

```{r}
train1 <- x[splitPlan[[1]]$train, ]

@@

# backwards step-wise selection with selectCox function
# converting to factors since function doesn't work with as.factor
train1$new_user <- as.factor(train1$new_user)
train1$num_freq_tags <- as.factor(train1$num_freq_tags)
train1$n_tags <- as.factor(train1$n_tags)

@@

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- MASS::stepAIC(first, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh, data = train1)
AIC(results2, k = 2) # 65277.47

@@

```{r}
# working with model from backward/forward stepwise selection

# adding interaction terms
# adding title_questionmark*title_beginwh
model2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(model2, k = 2) # 65274.49
compare_nested(model2, results2)

@@

* group all apple products together (use str_detect to search for Apple, ipad, ipod)
* recode left over Phones to Android/Other Phone
* group household + appliance as "Home"
* group car and truck + vehicle as "Vehicle"
* set a rule: any category with less than 50 questions gets grouped with "Other"

```{r}
# pie chart of product categories
ggplot(train1, aes(x = factor(1), fill = category)) + geom_bar(width = 1) + 
  coord_polar(theta = "y")

#===================================================
# recoding category variable
train1$new_category <- as.character(train1$category)

@@

# renaming left over phones 
train1$new_category[train1$new_category == "Phone"] <- "Android/Other Phone"
# merge household and appliance into "Home"
train1$new_category[train1$new_category == "Appliance" | train1$new_category == "Household"] <- "Home"
# merging Car and Truck with Vehicle 
train1$new_category[train1$new_category == "Car and Truck" | train1$new_category == "Vehicle"] <- "Vehicle"

# if there are less than 100 questions in a category, group them with other
counts <- train1 %>%
  group_by(new_category) %>%
  summarise(n = n())
for (i in which(counts$n <= 100)) {
  train1$new_category[train1$new_category == counts$new_category[i]] <- "Other"
}
table(train1$new_category)
#===================================================

@@

#new_cateogry*new_user
#new_cateogry*
model_new1 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_category*new_user, data = train1)
AIC(model_new1, k = 2) # 64994.49
compare_nested(model_new1, model_new)

@@

* Martingale residuals
    + plot for each continuous predictor 
    + assess adequacy of functional form of predictors, or suggest a potential form 
* Deviance residuals
    + check for outliers identify questions poorly predicted by the model 
    + one plot, examine questions with |residuals| > 2.5
* Schoenfeld residuals
    + one plot for each predictor 
    + if loess curve is flat, proportional hazard assumption is not violated, doesn't depend on time. Distinct pattern indicates violation of this assumption
    + **from the plots:** new_user, title_questionmark, num_freq_tags (1), ampm (Morning, Night), weekday (Thursday, Saturday), new_user*AMPM (morning, night) may violate PH assumption 
* Score residuals
    + check for influential observations. ID quesitons who's predictor values highly effect parameter estimates
    + plot for each predictor 
* **figure out how to identify certain points on ggplots**
* statified on new_category, AIC dropped drastically. But this means that I can't compare the different groups in new_category

```{r}
#=============================================
#martingale residuals
train1$mart <- residuals(model_full, type = "martingale")

@@

@@

# final model: stratified on new_category
final <- strat_model
```

@@

```{r}
train_predict <- predict(object = final, type = "risk", se.fit = TRUE)
train1$predictions <- train_predict[[1]]
performance <- coxph(Surv(time_until_answer, answered) ~ predictions, data = train1)
summary(performance)
AIC(performance, k = 2)

#using cph from rms package to get Dxy and other performance metrics
library(rms)
options(scipen=999)
(cox <- cph(Surv(time_until_answer, answered) ~ predictions, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE))
cox$stats
```

@@

```{r}

#calculating root

test1 <- x[splitPlan[[1]]$app, ]
# setting up variables for model 
test1 <- oshitar::variable_setup(test1)

test_predict <- predict(object = final, newdata = test1, type = "risk", se.fit = TRUE)
test1$predictions <- test_predict[[1]]

metric <- coxph(Surv(time_until_answer, answered) ~ predictions, data = test1)
summary(metric)
AIC(metric, k = 2)
```

```{r}
library(rms)
(cox <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE))

(cv <- validate(cox, method = "crossvalidation", B = 10, dxy = TRUE, pr = TRUE))
```

>>>>

```{r, data setup, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

@@

##### Fitting model with all significant variables (univariate analysis p-value < 0.01)

```{r, echo = FALSE}
train1 <- x[splitPlan[[1]]$train, ]

@@

# backwards step-wise selection with selectCox function
# converting to factors since function doesn't work with as.factor()
train1$new_user <- as.factor(train1$new_user)
train1$num_freq_tags <- as.factor(train1$num_freq_tags)
train1$n_tags <- as.factor(train1$n_tags)

@@

# backward/forward step-wise selection with stepAIC function from MASS package
model2 <- MASS::stepAIC(first, direction = "both", k = 2)
# results
results2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh, data = train1)
AIC(results2, k = 2) # 65277.47
```

@@

* working with model from backward/forward stepwise selection

```{r, echo = FALSE}
# adding title_questionmark*title_beginwh
model2 <- coxph(Surv(time_until_answer, answered) ~ category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(model2, k = 2) # 65274.49
compare_nested(model2, results2)

@@

* group all apple products together
    + use str_detect to search for ipad, ipod, apple watch 
    + group with iPhones and Macs 
* recode left over Phones to Android/Other Phone
* group household + appliance as "Home"
* group car and truck + vehicle as "Vehicle"
* set a rule: any category with less than 100 questions gets grouped with "Other"

```{r, echo = FALSE}
train1$new_category <- as.character(train1$category)

@@

# renaming left over phones 
train1$new_category[train1$new_category == "Phone"] <- "Android/Other Phone"

# merge household and appliance into "Home"
train1$new_category[train1$new_category == "Appliance" | train1$new_category == "Household"] <- "Home"

# merging Car and Truck with Vehicle 
train1$new_category[train1$new_category == "Car and Truck" | train1$new_category == "Vehicle"] <- "Vehicle"

# if there are less than 100 questions in a category, group them with Other
counts <- train1 %>%
  group_by(new_category) %>%
  summarise(n = n())
for (i in which(counts$n <= 100)) {
  train1$new_category[train1$new_category == counts$new_category[i]] <- "Other"
}
table(train1$new_category)

#===================================================

@@

#new_cateogry*new_user
model_new1 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + new_category*new_user, data = train1)
AIC(model_new1, k = 2) # 64994.49
compare_nested(model_new1, model_new)

@@

* Martingale residuals
    + plot for each continuous predictor 
    + assess adequacy of functional form of predictors, or suggest a potential form 
* Deviance residuals
    + check for outliers identify questions poorly predicted by the model 
    + one plot, examine questions with |residuals| > 2.5
* Schoenfeld residuals
    + one plot for each predictor 
    + if loess curve is flat, proportional hazard assumption is not violated, doesn't depend on time. Distinct pattern indicates violation of this assumption
    + **from the plots:** new_user, title_questionmark, num_freq_tags (1), ampm (Morning, Night), weekday (Thursday, Saturday), new_user*AMPM (morning, night) may violate PH assumption 
* Score residuals
    + check for influential observations. ID quesitons who's predictor values highly effect parameter estimates
    + plot for each predictor 
* **figure out how to identify certain points on ggplots**
* statified on new_category, AIC dropped drastically. But this means that I can't compare the different groups in new_category 
    + Concordance statistic and rsquare dropped

```{r, residuals, echo = FALSE}
#=============================================
#martingale residuals
train1$mart <- residuals(model_full, type = "martingale")

@@

#=============================================
# addression PH assumption violations

@@

# stratifying on ampm 
strat_model2 <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + strata(ampm) + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(strat_model2, k = 2) # 54447.8
summary(strat_model2)

# stratifying on ampm, new_user, new_category
strat_model3 <- coxph(Surv(time_until_answer, answered) ~ strata(new_category) + strata(new_user) + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + strata(ampm) + weekday + device_length + title_beginwh + title_questionmark*title_beginwh, data = train1)
AIC(strat_model3, k = 2) # 35370.81
summary(strat_model3)
```

@@

* model not stratified on new_category has higher R2, higher Dxy than model stratified on new_category
* survConcordance: 
    + Concordance is defined as Pr(agreement) for any two randomly chosen observations, where in this case agreement means that the observation with the shorter survival time of the two also has the larger risk score. The predictor (or risk score) will often be the result of a Cox model or other regression
    + some of the pairs are incomparable. For instance a pair of times (5+, 8), the first being a censored value. We do not know whether the first survival time is greater than or less than the second
    + observations that are comparable, pairs may also be tied on survival time (but only if both are uncensored) or on the predictor. The final concondance is (agree + tied/2)/(agree + disagree + tied)
    +

```{r}
# working with final model (stratified on new_category) 
train_predict <- predict(object = strat_model, type = "risk", se.fit = TRUE)
train1$predictions <- train_predict[[1]]

(performance <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE)) # Dxy 0.115
AIC(performance, k = 2) # 65781.26

survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train1) # Concordance = 0.5573537

Hmisc::rcorrcens(Surv(time_until_answer, answered) ~ predictions, data = train1)

#=============================================

# working with full_model (not stratified on new_category)
train_predict2 <- predict(object = model_full, type = "risk", se.fit = TRUE)
train1$predictions2 <- train_predict2[[1]]
(performance2 <- rms::cph(Surv(time_until_answer, answered) ~ predictions2, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE)) # Dxy 0.270

survConcordance(Surv(time_until_answer, answered) ~ predictions2, data = train1) # Concordance = 0.634903

Hmisc::rcorrcens(Surv(time_until_answer, answered) ~ predictions2, outx = TRUE, data = train1)

#=============================================

# working with strat_model1 (stratified on new_user)
train_predict3 <- predict(object = strat_model1, type = "risk", se.fit = TRUE)
train1$predictions3 <- train_predict3[[1]]
(performance3 <- rms::cph(Surv(time_until_answer, answered) ~ predictions3, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE)) # Dxy 0.244

survConcordance(Surv(time_until_answer, answered) ~ predictions3, data = train1) # Concordance = 0.6221597

Hmisc::rcorrcens(Surv(time_until_answer, answered) ~ predictions3, data = train1)

#=============================================

# working with strat_model2 (stratified on ampm)
train_predict4 <- predict(object = strat_model2, type = "risk", se.fit = TRUE)
train1$predictions4 <- train_predict4[[1]]
(performance4 <- rms::cph(Surv(time_until_answer, answered) ~ predictions4, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE)) # Dxy 0.258

survConcordance(Surv(time_until_answer, answered) ~ predictions4, data = train1) # Concordance = 0.6289692

Hmisc::rcorrcens(Surv(time_until_answer, answered) ~ predictions4, data = train1)
```

```{r, echo = FALSE}
(cox <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE))

options(scipen=999)
(cv <- rms::validate(cox, method = "crossvalidation", B = 10, dxy = TRUE, pr = TRUE))

# for calibrate function, had to add in surv = TRUE and time.inc = a constant (time at which to estimate survival probabilities)
cox1 <- rms::cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = train1, x = TRUE, y = TRUE, se.fit = TRUE, residuals = TRUE, surv=TRUE, time.inc=1.5)

calibrate <- rms::calibrate(cox1, method = "crossvalidation", B = 10, pr = TRUE)
```

@@

```{r, echo = FALSE}
# function to perform cross validation + compute performance metrics
crossvalidate <- function(train, test) {

  # building model on training data set
  model <- survival::coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = train)

  # performance metrics for training data 
  train[["predictions"]] <- predict(model, type = "risk")
  metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
  train_metrics <- data.frame(HR = exp(metric$coefficients), 
                              LR = round(metric$stats[3],2), 
                              pval = round(metric$stats[5],2), 
                              R2 = round(metric$stats[8],2), 
                              Dxy = round(metric$stats[9],2), 
                              AIC = stats::AIC(metric, k = 2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)

  # predicting on test data
  test[["predictions"]] <- predict(model, newdata = test, type = "risk")

  # computing performance metrics 
  metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
  test_metrics <- data.frame(HR = exp(metric1$coefficients), 
                             LR = round(metric1$stats[3],2), 
                             pval = round(metric1$stats[5],2), 
                             R2 = round(metric1$stats[8],2), 
                             Dxy = round(metric1$stats[9],2), 
                             AIC = stats::AIC(metric1, k = 2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)

  # returns data frame with train/test metrics 
  statistics <- rbind(train_metrics, test_metrics)
  rownames(statistics) <- c("Training Data", "Test Data")
  return(statistics) 
}

#=============================================
test1 <- oshitar::variable_setup(x[splitPlan[[1]]$app, ])
(iteration1 <- crossvalidate(train1, test1))

#=============================================
train2 <- oshitar::variable_setup(x[splitPlan[[2]]$train, ])
test2 <- oshitar::variable_setup(x[splitPlan[[2]]$app, ])
(iteration2 <- crossvalidate(train2, test2))

#=============================================
train3 <- oshitar::variable_setup(x[splitPlan[[3]]$train, ])
test3 <- oshitar::variable_setup(x[splitPlan[[3]]$app, ])
test3$num_freq_tags <- as.character(test3$num_freq_tags)
test3$num_freq_tags[test3$num_freq_tags == "4"] <- "3"
(iteration3 <- crossvalidate(train3, test3))
# make this more reproducible!!! rule: if a factors level has less than 3 observations in it, merge it with the nearest level

#=============================================
train4 <- oshitar::variable_setup(x[splitPlan[[4]]$train, ])
test4 <- oshitar::variable_setup(x[splitPlan[[4]]$app, ])
(iteration4 <- crossvalidate(train4, test4))

#=============================================
train5 <- oshitar::variable_setup(x[splitPlan[[5]]$train, ])
test5 <- oshitar::variable_setup(x[splitPlan[[5]]$app, ])
(iteration5 <- crossvalidate(train5, test5))

iteration1
iteration2
iteration3
iteration4
iteration5
```

# Fitting model to the full data set

```{r, echo = FALSE}
x <- oshitar::variable_setup(x)

final <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = x)
summary(final)
AIC(final, k = 2) # 83051.61

x$predictions <- predict(final, type = "risk")
(performance <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = x))
AIC(performance, k = 2)

# extracting predicted survival probabilities
prob <- pec::predictSurvProb(final, newdata = x, times = c(0.5, 1, 1.5, 2, quantile(x$time_until_answer)))

#==================================================================================================
# function to take in a data set, fit the model, output performance metrics, and allow you to specify times to predict survival probabilities

predict_survprob <- function(dataset, times) {
  # fitting the model 
  data <- oshitar::variable_setup(dataset)
  model <- coxph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + title_questionmark + num_freq_tags + contain_answered + text_all_lower + update + ampm + weekday + device_length + title_beginwh + title_questionmark*title_beginwh + new_user*update + new_user*ampm + new_category*new_user, data = data)

  # performance metrics
  data[["predictions"]] <- predict(model, type = "risk")
  p <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = data)
  metrics <- data.frame(HR = exp(p$coefficients), 
                             LR = round(p$stats[3],2), 
                             pval = round(p$stats[5],2), 
                             R2 = round(p$stats[8],2), 
                             Dxy = round(p$stats[9],2), 
                             AIC = stats::AIC(p, k = 2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = data)$concordance)

  # calculating predicted survival probabilities
  probs <- pec::predictSurvProb(model, newdata = data, times)

  # returns a list containing performance metrics, and data frame with predicted survival probabilities
  output <- list(metrics, probs)
  return(output)
}

```{r}
b <- oshitar::variable_setup(b)
counts <- b %>% 
  select(new_user, weekday, num_freq_tags) %>%
  purrr::map(~table(.))

# if a factor has less than 3 questions in a level, merge it with the nearest level 
for (i in length(counts)) {
  if (which(unname(counts[[i]]) < 3) != integer(0)) {
    names(which(counts[[i]] < 3)) 
  }
}

class(counts[[1]])
which(unname(counts[[1]]) < 1)
```
